Fully Automatic Loop-free Program Synthesis by Bounded Model Checking

### 2014-01-28 paper/paper.tex
Space saving measures

\usepackage{listings}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{pgf}
\usepackage[noend]{algpseudocode}
\usepackage{algorithm}
\usepackage{amsfonts}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{pifont}

@@

 \begin{algorithmic}[1]
\Statex
\Function{synth}{inputs}
  \Let{$(i_1, \ldots, i_N)$}{inputs}
  \Let{query}{$\exists P . \sigma(i_1, P(i_1)) \land \ldots \land \sigma(i_N, P(i_N))$}
  \Let{result}{decide(query)}
  \If{result.satisfiable}
    \State \Return{result.model}
  \Else
    \State \Return{unsatisfiable}
  \EndIf
\EndFunction
\Statex
\Function{verif}{P}
  \Let{query}{$\exists x . \lnot \sigma(x, P(x))$}
  \Let{result}{decide(query)}
  \If{result.satisfiable}
    \State \Return{result.model}
  \Else
    \State \Return{valid}
  \EndIf
\EndFunction
\Statex
\Function{refinement loop}{}
  \Let{inputs}{$\emptyset$}
  \Loop
    \Let{candidate}{\Call{synth}{inputs}}
    \If{candidate = unsatisfiable}
      \State \Return{unsatisfiable}
    \EndIf
    \Let{res}{\Call{verif}{candidate}}
    \If{res = valid}
      \State \Return{candidate}
    \Else
      \Let{inputs}{inputs $\cup$ res}
    \EndIf
  \EndLoop
\EndFunction
 \end{algorithmic}
\end{algorithm}

\begin{figure}
 \centering
 \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,
 semithick, initial text=]

@@

To synthesise tricky bitvector programs like this, it is natural for us to
work in the logic of quantifier-free propositional formulae and to use a
propositional SAT or SMT-$\mathcal{BV}$ solver as the decision procedure. 
However we propose a slightly different tack, which is to use a decidable
fragment as C as a ``high level'' logic.

@@

\begin{figure}
\begin{center}
\setlength{\tabcolsep}{16pt}
Integer arithmetic instructions:

@@

\begin{tabular}{llll}
 \verb|le  a b| & \verb|lt  a b| & \verb|sle  a b| & \verb|slt a b|
\end{tabular}
\end{center}

 \caption{The language $\mathcal{L}$}
 \label{fig:l-language}

@@

\begin{figure}
\centering
\begin{minipage}[t]{0.43\textwidth}
\begin{lstlisting}[language=C++]
void synth() {
  prog_t p = nondet();
  int in[N], out[M];

  in = test1;
  exec(p, in, out);
  assume(check(in, out));

  ...

  in = testN;
  exec(p, in, out);
  assume(check(in, out));

  assert(false);
}
\end{lstlisting}
\end{minipage}
\vrule \quad
\begin{minipage}[t]{0.45\textwidth}
\begin{lstlisting}[language=C]
void verif(prog_t p) {
  int in[N] = nondet();
  int out[M];

@@

\begin{tikzpicture}[font=\sffamily]

@@

\end{tikzpicture}
\end{center}

@@

The exact \newC encoding of an $\mathcal{L}$ program is shown in Fig.~\ref{fig:c-l-encoding}.
The \verb|prog_t| structure encodes a complete program, which is a sequence of instructions.
The parameter $a$ is the number of arguments the program takes.
The $i$th instruction has opcode \verb|ops[i]|, left operand \verb|params[i*2]| and
right operand \verb|params[i*2 + 1]|.  An operand refers to either a program constant,
a program argument or the result of a previous instruction.  The value of an operand
is determined at runtime as follows:

\[
 val(x) = \begin{cases}
           x < a & \text{the } x^{\text{th}} \text{ program argument} \\
           a \leq x < a+c & \mathtt{consts[} x-a \mathtt{]} \\
           x \geq a + c & \text{the result of the } (x - a - c)^{\text{th}} \text{ instruction}
          \end{cases}
\]

@@

\begin{figure}
\begin{tikzpicture}[scale=0.9, transform shape, node distance=2cm, auto]
\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
    text width=4.5em, text badly centered, node distance=3cm, inner sep=0pt]
\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
    text width=5em, text centered, rounded corners, minimum height=4em]
\tikzstyle{line} = [draw, -latex']
\tikzstyle{cloud} = [draw, ellipse,fill=red!20, node distance=3cm,
    minimum height=2em]

@@

 \node [block, below of=verif, node distance=2.5cm]   (done) {Done!};
 \node [decision, left of=done] (verifw) {{\sc Verif} succeeds for small words?};

@@

 \node [decision, below of=verifw, node distance=3cm] (gen) {Extend succeeds?};
 \node [block, left of=gen, node distance=2.5cm] (iterate) {Parameters unchanged};

 \node [block, below of=gen, node distance=2.5cm] (incw) {$w := w+1$};

@@

 \path [line] (verif) -- node [left] {Yes} (done);
 \path [line] (verif) -| node [above, near start] {No} (verifw);

@@

 \path [line] (verifw) -- node [right] {Yes} (gen);
 \path [line] (verifw) -| node [above] {No} (iterate);

 \path [line] (gen) -| node [below, near start] {Yes} (done);
 \path [line] (gen) -- node [right] {No} (incw);

@@

\begin{figure}
\begin{center}
\begin{minipage}[t]{0.45\linewidth}
\begin{algorithm}[H]
\caption{\sc 2Sum
 \label{alg:2sum}}
\begin{algorithmic}
\Let{$s$}{$a+b$}
\Let{$b'$}{$s - a$}
\Let{$a'$}{$s - b'$}
\Let{$\delta b$}{$b - b'$}
\Let{$\delta a$}{$a - a'$}
\Let{$t$}{$\delta a + \delta b$}
\Ensure{$a + b$ = $s + t$ exactly}
\end{algorithmic}
\end{algorithm}
\end{minipage}
\hfill
\begin{minipage}[t]{0.45\linewidth}
\begin{algorithm}[H]
\caption{\sc Fast2Sum
 \label{alg:fast2sum}}
\begin{algorithmic}
\Require{$a \ge b$}
\Let{$s$}{$a + b$}
\Let{$z$}{$s - a$}
\Let{$t$}{$b - z$}
\Ensure{$a + b$ = $s + t$ exactly}
\end{algorithmic}
\end{algorithm}
\end{minipage}
\end{center}

 \caption{The {\sc 2Sum} algorithm for correctly rounded floating-point sums.}
  \label{fig:2sum}
\end{figure}

\begin{figure}[p]
\begin{center}
{\small
\include{figures/results_table}
}
\end{center}
\caption{Time for synthesis of machine integer benchmarks from~\cite{brahma} and~\cite{brahma-icse}.
 \label{fig:results-table}}
\end{figure}

@@

\begin{figure}[p]
{\tiny
\begin{center}
 \begin{tabular}{r||r|r|r|r|r|r|r|r|r}
  & \multicolumn{3}{c}{Count} & \multicolumn{2}{|c}{Total time} & \multicolumn{2}{|c}{Absolute speedup} & \multicolumn{2}{|c}{Relative speedup} \\
  {\sc Brahma} version & Total & {\sc Kalashnikov} & {\sc Brahma} & {\sc Kalashnikov} & {\sc Brahma} & Mean & Median & Mean & Median \\
  &       & faster            & faster       &                   &              &      &        &      & \\
  \hline
  \hline
  PLDI & 16 & 13 & 3 & 184.12\,s & 544.20\,s & 22.50\,s & 2.11\,s & 2.60 & 1.77 \\
  ICSE Semibiased & 17 & 7 & 10 & 225.49\,s & 262.24\,s & 2.16\,s & -1.00\,s & 1.26 & 0.87
 \end{tabular}
\end{center}
}

 \caption{Performance statistics when {\sc Brahma} and {\sc Kalashnikov} were both successful.
 \label{fig:stats-table}}
\end{figure}

>>>>

\usepackage{listings}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{pgf}
\usepackage[noend]{algpseudocode}
\usepackage{algorithm}
\usepackage{amsfonts}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{pifont}
\usepackage{multicol}

@@

 \begin{multicols}{2}
 \begin{algorithmic}[1]
\Statex
\Function{synth}{inputs}
  \Let{$(i_1, \ldots, i_N)$}{inputs}
  \Let{query}{$\exists P . \sigma(i_1, P(i_1)) \land \ldots \land \sigma(i_N, P(i_N))$}
  \Let{result}{decide(query)}
  \If{result.satisfiable}
    \State \Return{result.model}
  \Else
    \State \Return{unsatisfiable}
  \EndIf
\EndFunction
\Statex
\Function{verif}{P}
  \Let{query}{$\exists x . \lnot \sigma(x, P(x))$}
  \Let{result}{decide(query)}
  \If{result.satisfiable}
    \State \Return{result.model}
  \Else
    \State \Return{valid}
  \EndIf
\EndFunction
\columnbreak
\Statex
\Function{refinement loop}{}
  \Let{inputs}{$\emptyset$}
  \Loop
    \Let{candidate}{\Call{synth}{inputs}}
    \If{candidate = UNSAT}
      \State \Return{unsatisfiable}
    \EndIf
    \Let{res}{\Call{verif}{candidate}}
    \If{res = valid}
      \State \Return{candidate}
    \Else
      \Let{inputs}{inputs $\cup$ res}
    \EndIf
  \EndLoop
\EndFunction
 \end{algorithmic}
 \end{multicols}
\end{algorithm}

\begin{figure}
 \centering
 \begin{tikzpicture}[scale=0.5,->,>=stealth',shorten >=1pt,auto,
 semithick, initial text=]

@@

To synthesise tricky bitvector programs like this, it is natural for us to
work in the logic of quantifier-free propositional formulae and to use a
propositional SAT or SMT-$\mathcal{BV}$ solver as the decision procedure. 
However we propose a slightly different tack, which is to use a decidable
fragment of C as a ``high level'' logic.

@@

\begin{figure}[h]
{\small
\begin{center}
\setlength{\tabcolsep}{16pt}
Integer arithmetic instructions:

@@

\begin{tabular}{llll}
 \verb|le  a b| & \verb|lt  a b| & \verb|sle  a b| & \verb|slt a b|
\end{tabular}
\end{center}
}
 \caption{The language $\mathcal{L}$}
 \label{fig:l-language}
\end{figure}

@@

\begin{figure}[h]
\centering
\begin{minipage}[t]{.45\textwidth}
\begin{lstlisting}[language=C++]
void synth() {
  prog_t p = nondet();
  int in[N], out[M];

  assume(wellformed(p));

  in = test1;
  exec(p, in, out);
  assume(check(in, out));
  ...
  in = testN;
  exec(p, in, out);
  assume(check(in, out));

  assert(false);
}
\end{lstlisting}
\end{minipage}
\hfill
\begin{minipage}[t]{.45\linewidth}
\begin{lstlisting}[language=C++]
void verif(prog_t p) {
  int in[N] = nondet();
  int out[M];

@@

\resizebox{!}{2.5cm}{
\begin{tikzpicture}[font=\sffamily]

@@

\end{tikzpicture}
}
\end{center}

@@

The exact \newC encoding of an $\mathcal{L}$ program is shown in Fig.~\ref{fig:c-l-encoding}.
The \verb|prog_t| structure encodes a program, which is a sequence of instructions.
The parameter $a$ is the number of arguments the program takes.
The $i$th instruction has opcode \verb|ops[i]|, left operand \verb|params[i*2]| and
right operand \verb|params[i*2 + 1]|.  An operand refers to either a program constant,
a program argument or the result of a previous instruction, and its value
is determined at runtime as follows:
\[
 val(x) = \begin{cases}
           x < a & \text{the } x^{\text{th}} \text{ program argument} \\
           a \leq x < a+c & \mathtt{consts[} x-a \mathtt{]} \\
           x \geq a + c & \text{the result of the } (x - a - c)^{\text{th}} \text{ instruction}
          \end{cases}
\]

@@

\begin{figure}
\centering
\begin{tikzpicture}[scale=0.7, transform shape, node distance=2cm, auto]
\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
    text width=4.5em, text badly centered, node distance=3cm, inner sep=0pt]
\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
    text width=5em, text centered, rounded corners, minimum height=4em]
\tikzstyle{line} = [draw, -latex']
\tikzstyle{cloud} = [draw, ellipse,fill=red!20, node distance=3cm,
    minimum height=2em]

@@

 \node [block, below of=verif, node distance=3cm]   (done) {Done!};
 \node [decision, left of=verif] (verifw) {{\sc Verif} succeeds for small words?};

@@

 \node [decision, below of=verifw, node distance=3cm] (gen) {Extend succeeds?};
 \node [block, left of=verifw, node distance=3cm] (iterate) {Parameters unchanged};

 \node [block, left of=gen, node distance=3cm] (incw) {$w := w+1$};

@@

 \path [line] (verif) -- node [left] {Yes} (done);
 \path [line] (verif) -- node [above, near start] {No} (verifw);

@@

 \path [line] (verifw) -- node [right] {Yes} (gen);
 \path [line] (verifw) -- node [above] {No} (iterate);

 \path [line] (gen) -- node [below] {Yes} (done);
 \path [line] (gen) -- node [below] {No} (incw);

@@

\iffalse
\begin{figure}
\begin{center}
\begin{minipage}[t]{0.45\linewidth}
\begin{algorithm}[H]
\caption{\sc 2Sum
 \label{alg:2sum}}
\begin{algorithmic}
\Let{$s$}{$a+b$}
\Let{$b'$}{$s - a$}
\Let{$a'$}{$s - b'$}
\Let{$\delta b$}{$b - b'$}
\Let{$\delta a$}{$a - a'$}
\Let{$t$}{$\delta a + \delta b$}
\Ensure{$a + b$ = $s + t$ exactly}
\end{algorithmic}
\end{algorithm}
\end{minipage}
\hfill
\begin{minipage}[t]{0.45\linewidth}
\begin{algorithm}[H]
\caption{\sc Fast2Sum
 \label{alg:fast2sum}}
\begin{algorithmic}
\Require{$a \ge b$}
\Let{$s$}{$a + b$}
\Let{$z$}{$s - a$}
\Let{$t$}{$b - z$}
\Ensure{$a + b$ = $s + t$ exactly}
\end{algorithmic}
\end{algorithm}
\end{minipage}
\end{center}

 \caption{The {\sc 2Sum} algorithm for correctly rounded floating-point sums.}
  \label{fig:2sum}
\end{figure}
\fi

\begin{figure}
\begin{center}
{\small
\include{figures/results_table}
}
\end{center}
\caption{Time for synthesis of machine integer benchmarks from~\cite{brahma} and~\cite{brahma-icse}.
 \label{fig:results-table}}
\end{figure}

@@

\begin{figure}
{\small
 \include{figures/stats_table}
}
 \caption{Performance statistics when {\sc Brahma} and {\sc Kalashnikov} were both successful.
 \label{fig:stats-table}}
\end{figure}
